Value-Suppressing Uncertainty Maps

### 2017-03-22 paper/figures.tex
Added uncertainty references
Submitted abstract version.

\newcommand{\teaserFig}{
  \teaser{
		\centering
		\includegraphics[width=1.5in]{sample.eps}
		\caption{Lookit! Lookit!}
		\label{fig:teaser}
	}
}

>>>>

%Figure list:
% Vsum vs. traditional 2D heatmap vs. juxtaposed maps
% Chart of color bins w/r/t CIELAB threshold, with iconic maps at intervals
% Process figure
% Real examples with different color maps

\newcommand{\teaserFig}{
  \teaser{
		\centering
		\includegraphics[width=0.9\textwidth]{example.pdf}
		\caption{Lookit! Lookit!}
		\label{fig:teaser}
	}
}

\newcommand{\exampleFig}{
\begin{figure}
	\centering
	\includegraphics[width=0.9\columnwidth]{example.pdf}
	\caption{A comparison between a standard bivariate map and a VSUM. Both use the same visual channels to encode value (position along the Viridis color map) and uncertainty (lightness and saturation), and have an equal standard of perceptual discriminability (at least 18 units of distance in CIELAB color space between colors). However, highly uncertain values result in colors that are very close together in the bivariate case, meaning the full bivariate map is only 9 bins under these constraints--- a 4x4 map would result in colors that are perceptually too close together. By contrast, the VSUM intentionally reduces bins when uncertainty is high, eventually aliasing all highly uncertain values to the same color. This decision affords more distinct colors in other regions of the map, and so an increase in overall bins to 15. The resulting map suppresses the value of uncertain data, but increases the discriminability of highly certain data.}
	\label{fig:example}
\end{figure}

}

### 2017-03-22 paper/template.tex
Submitted abstract version.

Uncertainty is a vital concern in many datasets. Visualizations often use bivariate mappings to encode value and uncertainty simultaneously. Due to interference between visual channels, these bivariate maps are often limited in resolution. We contribute Value-Suppressing Uncertainty Maps (VSUMs), an encoding technique that intentionally impairs value discrimination as uncertainty increases. In contrast to traditional bivariate maps, VSUMs alias values when uncertainty is high: expected values with lower uncertainty can be more precisely compared, whereas similar values with higher uncertainty are made harder to differentiate. By making informed binning decisions, VSUMs seek to better use the limited budget of discriminable marks in bivariate maps, and dissuade analysts from making decisions based on data with high uncertainty. We demonstrate several examples of VSUMs and present a crowdsourced evaluation showing that, compared to traditional bivariate maps, VSUMs bias analysts to more heavily weight uncertainty information in decision-making tasks.

@@

Uncertainty is an inescapable component of collecting, presenting, and using data. A common goal in the communication of uncertainty is promoting \textbf{uncertainty-aware decisions}--- that is, the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty. In general, a goal of designers of visualizations of uncertain data is the \emph{information fusion} of data and uncertainty~\cite{riveiro2007evaluation}. A way that designers can contribute to this goal is by ensuring that uncertainty information is \textbf{well-integrated} with the rest of the data. That is, it should be difficult to discount or ignore the uncertainty in a dataset. Simultaneous presentation of uncertainty and value necessitates the construction of a bivariate map--- a relation, in terms of visual variables, between a 2-tuple $(value$,$uncertainty)$ and a set of marks.

In this paper, we present \textbf{Value-Suppressing Uncertainty Maps} (VSUMs) for integrating data and uncertainty information in thematic maps. VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as certainty increases. Rather than a traditional two dimensional bivariate map, with unique outputs for each relevant combination of data and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value. Figure X shows an example of a VSUM, compared to a more traditional bivariate map. In this work, we describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and validate that VSUMs in a crowd-sourced experiment.

\section{Related Works}

% What this section needs to do:
% State of the art in uncertainty vis, with heavy emphasis on MacEachren & co. 
% Bivariate maps are hard!
% Binning colormaps is de rigueur
% The variables we'd want to use for uncertainty (like value or alpha or size) mess with color discriminability
Despite the acknowledged importance of uncertainty in understanding data, explicit representation of uncertainty often missing from visualizations~\cite{boukhelifa2009uncertainty}. The absence of

>>>>

Uncertainty is a vital concern in many datasets. Visualizations often use bivariate mappings to encode value and uncertainty simultaneously. Due to interference between visual channels, these bivariate maps are often limited in discriminability. We contribute Value-Suppressing Uncertainty Maps (VSUMs), an encoding technique that intentionally impairs value discrimination as uncertainty increases. In contrast to traditional bivariate maps, VSUMs alias values when uncertainty is high: expected values with lower uncertainty can be more precisely compared, whereas similar values with higher uncertainty are made harder to differentiate. By making informed binning decisions, VSUMs seek to better use the limited budget of discriminable marks in bivariate maps, and dissuade analysts from making decisions based on data with high uncertainty. We demonstrate several examples of VSUMs and present a crowdsourced evaluation showing that, compared to traditional bivariate maps, VSUMs bias analysts to more heavily weight uncertainty information in decision-making tasks.

@@

Uncertainty is an inescapable component of collecting, presenting, and using data. A common goal in the communication of uncertainty is promoting \textbf{uncertainty-aware decisions}--- that is, the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty.  A way that designers can contribute to this goal is by ensuring that uncertainty information is \textbf{well-integrated} with the rest of the data. That is, it should be difficult to discount or ignore the uncertainty in a dataset. Simultaneous presentation of uncertainty and value necessitates the construction of a bivariate map--- a relation, in terms of visual variables, between a 2-tuple $(value$,$uncertainty)$ and a set of marks. The design of bivariate maps is difficult and, due to the interference and interplay between different visual variables, often leads to maps with very few discriminable categories.

In this paper, we present \textbf{Value-Suppressing Uncertainty Maps} (VSUMs) for integrating data and uncertainty information in thematic maps. VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as certainty increases. Rather than a traditional two dimensional bivariate map, with unique outputs for each relevant combination of data and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value. Figure \ref{fig:example} shows an example of a VSUM, compared to a more traditional bivariate map. In this work, we describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and validate that VSUMs in a crowd-sourced experiment.

\exampleFig 
\section{Related Works}

% What this section needs to do:
% State of the art in uncertainty vis, with heavy emphasis on MacEachren & co. 
% Bivariate maps are hard!
% Binning colormaps is de rigueur
% The variables we'd want to use for uncertainty (like value or alpha or size) mess with color discriminability

Despite the acknowledged importance of uncertainty in understanding data, explicit representation of uncertainty often missing from visualizations~\cite{boukhelifa2009uncertainty}. The absence of
